### 2013-12-15 paper/paper.tex
Add citations

\documentclass{article}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{fancyhdr}

\usepackage{hyperref}
\setlength{\headheight}{15pt}
\pagestyle{fancyplain}

\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex}

\tikzstyle{int}=[draw, minimum size=4em]
\tikzstyle{init} = [pin edge={to-,thin,black}]
\lhead{Dynamic Hover Detection: Creating GUIs in Elm}
\rhead{Goldstein \& Alter}

\makeatletter
\renewcommand{\maketitle}{ % Customize the title - do not edit title and author name here, see the TITLE block below
    \begin{flushright} % Right align
        {\LARGE\@title} % Increase the font size of the title

        \vspace{40pt} % Some vertical space between the title and author name

        {\large\@author} % Author name

        \vspace{30pt} % Some vertical space between the author block and abstract
    \end{flushright}
}

% This begins the document
\title{\textbf{Dynamic Hover Detection:\\Creating GUIs in Elm}}
\author{Max Goldstein \& Eliot Alter}

\begin{document}
\thispagestyle{empty} % No header on first page
\maketitle

\begin{abstract}
The language Elm was designed to be a practical application of
Functional Reactive Programming (FRP) to create Graphical User
Interfaces (GUIs). We attempted to create a menu bar as commonly seen on
desktop operating systems, and quickly discovered an apparent
incompatibility between Elm's mouse-hover detection primitive and the
constraints of the language. This concern only surfaces when dealing
with highly dynamic data. We present a non-obvious technique to address
the issue without modifying the Elm compiler. It generalizes to other
functions in Elm's \texttt{Graphics.Input} library, which includes
(besides hover detection) GUI mainstays such as buttons, checkboxes, and
text fields. We also contribute other techniques for the representation and
display of menus in Elm, and contrast our work with an existing Elm web
application.
\end{abstract}

\section{Introduction}\label{introduction}

Elm was introduced in March 2012 in Evan Czaplicki's senior thesis,
\emph{Concurrent FRP for GUIs}. \cite{elm-thesis} He and his adviser, Stephen
Chong, published \emph{Asynchronous FRP for GUIs} at PLDI
2013. \cite{CzaplickiC13} Both papers are comprehensive overviews of Elm, and
additionally provide excellent literature reviews of previous FRP GUI endeavors,
of which there are several.  Elm compiles down to JavaScript to run in the
browser, making it remarkably portable.

We chose menus as an example GUI to implement in Elm. The particular
design places the top-level menu at the top of the screen, similar to
Mac OS X and many Linux distributions, with selections coming down from
the top. Menus extend when the top-level item is hovered upon, and
remain extended while the mouse hovers over any item in the menu.
Therefore it is necessary to know hover information about each menu
item. This time-varying information is also used to detect selections
upon click and highlight the moused-over item. It is simple to do this
when the hover-detecting area is constant. Our paper describes the much
more difficult task of managing time-varying hover information about
time-varying areas.

There are two features of Elm we are deliberately avoiding. First is the
extensive raster drawing library, \texttt{Graphics.Collage}. Dynamic hover
detection is not problematic when using this library because it can be done
purely through manual collision detection. However, if we used this library, our
GUI would be a single raster animation and not a DOM tree. Secondly, the
\texttt{Graphics.Input} library contains wrappers around HTML checkboxes and
dropdowns. We rejected creating GUIs with these and wanted to pursue something
more general. This led to the choice of an interface typically found in the
operating system rather than the browser. We do refer to the native GUI
constructs to show how the technique we develop for hovering generalizes, as the
API for buttons and hover detection is very similar.

It is difficult for the authors to assess what level of knowledge should
be assumed on the part of the reader. Firstly, readers will range from
Elm's creators and experienced users, who are already familiar with its
restrictions and standard libraries, to those with no FRP background,
who need an introduction to signals. Secondly, while we have found no
prior discussion on either the problem we identify nor its solution, we
cannot know for certain that either are novel. We continue in the hopes
of presenting new and non-trivial techniques to the Elm community.

We contribute:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The identification of an apparent limitation in Elm's hover detection
  library, solved by a non-trivial usage pattern that does not require
  modifications to the Elm compiler or runtime, and that generalizes to
  other functions in \texttt{Graphics.Input}.
\item
  An implementation of desktop-style menu in Elm, which incorporates
  several noteworthy ``tricks'' as well as general style practices.
\item
  An analysis of TodoFRP, the current state-of-the-art in dynamic Elm
  GUIs. We demonstrate how it operates in the absence of our technique,
  and how it could operate in its presence.
\end{itemize}

Section \ref{elm-for-functional-programmers} introduces Elm, signals, state, and
the prohibition on signals of signals. It is targeted to readers familiar with
functional programming but not FRP, and may be skipped by those already
comfortable with Elm.
Section \ref{detecting-hover-information} presents Elm's hover detection for DOM
elements, why it initially appears inadequate, and how it can be extended to
meet our needs.
Section \ref{implementing-menus} explains our menu implementation in detail.
Section \ref{related-work-todofrp} analyzes TodoFRP.
Section \ref{conclusion} concludes with a notice to the Elm community.

\section{Elm for Functional
Programmers}\label{elm-for-functional-programmers}

A typical functional program is \emph{transformative}: all input is
available at the start of execution, and after a hopefully finite amount
of time the program terminates with some output. In contrast, Elm
programs are \emph{reactive}: not all input is available immediately and
the program may indefinitely adjust output with each input. In simple
programs, the inputs at a given time fully determine the output. More
complex programs will take advantage of Elm's ability to remember state.

\subsection{Signals: Time-varying
values}\label{signals-time-varying-values}

A time-varying value of a polymorphic type \texttt{a} is represented by
\texttt{Signal a}. For example, the term \texttt{constant 150} has type
\texttt{Signal Int}. The combinator \texttt{constant} creates a signal
whose value never changes. A more interesting signal is the primitive
\texttt{Window.dimensions : Signal (Int, Int)}. This signal represents
the browser window size and updates whenever it is resized. Signals are
asynchronous in that they update at no set time, just as the window may
remain the same size indefinitely. Signals update in discrete events, but
are continuous in the sense that they are always defined.

The function \texttt{lift} allows us to execute a pure function on a
signal of inputs, producing a signal of outputs. (Although lifting is a
general functional concept, in Elm it has only this meaning.)

\begin{verbatim}
lift : (a -> b) -> Signal a -> Signal b
\end{verbatim}

For example, we can multiply the width and height of the window together
to find its area.

\begin{verbatim}
area : Signal Int
area = lift (uncurry (*)) Window.dimensions
\end{verbatim}

In this case, the lifted function is multiplication, uncurried as to
operate on pairs. As the Window library exposes width and height as both
a pair and individually, we can also write
\texttt{area2 = lift2 (*) Window.width Window.height}.

We can print the current area to the screen with \texttt{main = lift asText
area}. The primitive \texttt{asText : a -\textgreater{} Element} renders almost
anything into an Element, which represents a DOM element.\footnote{Those
    following along in an Elm compiler, such as the one available at
    \texttt{elm-lang.org/try}, should add \texttt{import Window} to the top of
the file.}

The signals in an Elm program can be thought of as a Directed Acyclic
Graph (DAG). Many signals depend on other signals for their output. For
example, the \texttt{area} signal depends on \texttt{Window.dimensions}
signal. Similarly, the \texttt{area2} signal depends on both the values
of the \texttt{Window.width} and the \texttt{Window.height} signals.
When an event is fired on a signal, it propagates down the DAG to all
signals who depend on that signal and the outputs of those signals are
reevaluated.

\subsection{Remembering State}\label{remembering-state}

Signals can remember state by using the \texttt{foldp} combinator.
Familiar list folds apply a binary operation of an element and an
accumulator to produce a new accumulator, over each list element in
sequence. Folding from the \emph{past} operates on all of the values of
a signal as they occur and produces a signal of the accumulator.

\begin{verbatim}
foldr : (a -> b -> b) -> b -> [a]      -> b
foldp : (a -> b -> b) -> b -> Signal a -> Signal b
\end{verbatim}

When the event signal updates, a pure function is called with the new
event and the old accumulator (a default is supplied), producing a new
accumulator that is the new value of the output signal. For example,
\texttt{foldp max 0 Window.width} is a signal of the maximum width ever
obtained by the window. With \texttt{foldp}, it is possible to create
signals that depend on every event to ever occur on a signal. However,
most folded functions do not store every value explicitly (cons is an
exception) and space can be saved by remembering only the accumulator.

\subsection{No Signals of Signals}\label{no-signals-of-signals}

The ability to create a signal dependent on past state has a potentially
disastrous implication for space performance. Czaplicki and Chong
explain the preventive measure and the problem:

\begin{quote}
The type system rules out programs that use signals of signals, for the
following reason. Intuitively, if we have signals of signals, then after
a program has executed for, say 10 minutes, we might create a signal
that (through the use of \texttt{foldp}) depends on the history of an
input signal, say \texttt{Window.width}. \cite{CzaplickiC13}
\end{quote}

That is, we can define a clock that measures the amount of time since
program execution began:

\begin{verbatim}
clock : Signal Time
clock = foldp (+) 0 (fps 10)
\end{verbatim}

The \texttt{fps n} combinator produces a signal of the time elapsed
since its last event, updated \texttt{n} times a second. We sum these
deltas starting with zero. Then we create a function from a time to one
of two signals, one of which is trivial and one of which depends on the
history of \texttt{Window.width}:

\begin{verbatim}
switcher : Time -> Signal Int
switcher t = if t < 10*minute
             then constant 0
             else foldp max 0 Window.width
\end{verbatim}

We could, were the type system not principled on disallowing it, create

\begin{verbatim}
switched : Signal (Signal Int)
switched = lift switched clock
\end{verbatim}

Why is this problematic? Czaplicki and Chong continue,

\begin{quote}
To compute the current value of this signal, should we use the entire
history of \texttt{Window.width}? But that would require saving all
history of \texttt{Window.width} from the beginning of execution, even
though we do not know whether the history will be needed later.
Alternatively, we could compute the current value of the signal just
using the current and new values of \texttt{Window.width} (i.e.,
ignoring the history). But this would allow the possibility of having
two identically defined signals that have different values, based on
when they were created. We avoid these issues by ruling out signals of
signals. \cite{CzaplickiC13}
\end{quote}

That is, in order to have both state and signals of signals, the program
must either remember indefinitely every event to ever occur, so that a
newly created signal can use them, or tolerate signals that vary based
only on when they were created, losing referential transparency. Elm's
creators declined either option and disallowed signals of signals.

To a reader familiar with Haskell, this means signals are functors (and
in fact applicative functors) but not monads, as monads support the
following operation:

\begin{verbatim}
join :: Monad m => m (m a) -> m a
\end{verbatim}

Such an operation for signals would condense a
\texttt{Signal (Signal a)} into a mere \texttt{Signal a}, but it cannot
exist in general.

\section{Detecting Hover Information}\label{detecting-hover-information}
A cursory inspection of \texttt{Graphics.Input} leads to the conclusion that
menus cannot be implemented as we had hoped. The remarkable solution generalizes
from detecting hover information to other GUI functions in the library.

\subsection{A Na\"{i}ve Menu}\label{a-naive-menu}

Elm's \texttt{Graphics.Input} library provides a function to obtain the hover
information from an Element.

\texttt{hoverable : Element -\textgreater{} (Element, Signal Bool)}

The returned Element is visually identical to argument, but now detects
hover information. The signal of Booleans reflects the hover status of
the this Element, not the original. This function works well when the
Element is pure (not a signal). For example, if the top-level menu items are
known statically and never change, they can be rendered as pure Elements with
this function. We can also create a function, perhaps of type \texttt{Bool ->
Element}, which we can lift onto the hover data to display the submenu only on
mouseover. The result is a value of type \texttt{Signal Element}. If we try to
lift \texttt{hoverable} to accept this value, we get

\texttt{lift hoverable : Signal Element -\textgreater{} Signal (Element, Signal Bool)}

Although it may not appear to be so at first, the result type is a signal of
signals. It is possible to transform the result into \texttt{Signal Signal
(Element, Bool)}. A more mentally convenient type, though impossible to
obtain,\footnote{We explain why in Section \ref{implementing-menus}.}
is \texttt{(Signal Element, Signal Signal Bool)}. There does not exist a general
join function to operate on the \texttt{Signal Signal Bool}.

This is rather unfortunate, as we need to detect hover information on a dynamic
element for a number of reasons. First, after a menu is extended by hovering
over its parent, it must remain so as long as the mouse hovers over the menu
itself. Without this information, the menu will disappear as soon as we try to
mouse over it. Secondly, if we wish the indicate the currently hovered-over item
to the user by visually highlighting it, the element must be dynamic. Thirdly,
we do not wish to require that menu labels be string literals in the source
code, which would be necessary if we could not use signals of elements. Rather,
they may be retrieved or generated in some other manner, or they may change as
the user switches among applications. Finally, in order to determine when a user
has clicked on an Element so that the system may respond, we must know which one
is being hovered over.\footnote{Once this is known, \texttt{sampleOn
Mouse.clicks} can be applied to obtain a signal of the hovered-over element
updated on each click.}

\subsection{Dynamic Hovering}
However, it is possible to implement the following function:

\texttt{hoverableJoin: Signal Element -\textgreater{} (Signal Element, Signal Bool)}

Comparing its type to the mentally convenient type of \texttt{lift hoverables},
we see it is identical except that the Boolean value is a signal, not a signal
of signals. This function is implemented using the more general
\texttt{hoverables} (note the plural) primitive, of the following type:

\begin{verbatim}
hoverables : a -> { events : Signal a,
                    hoverable : (Bool -> a) -> Element -> Element }
\end{verbatim}

The polymorphic \texttt{a} type can serve as an identifier. The first
value supplies the default value of \texttt{events} (signals must always
be defined and so a default value is required). The returned record
includes the \texttt{events} signals and the \texttt{hoverable}
function, which in general may be applied multiple times so that
multiple elements report on \texttt{events}. The
\texttt{(Bool -\textgreater{} a)} is used to identify which Element
experienced the event; a common use is \texttt{a = (Int, Bool)} where
the integer identifies the Element and the Bool is the event.
Additionally, \texttt{hoverables} is used to implement
\texttt{hoverable}:

\begin{verbatim}
hoverable : Element -> (Element, Signal Bool)
hoverable elem =
    let pool = hoverables False
        in  (pool.hoverable id elem, pool.events)
\end{verbatim}

It ignores the polymorphism (\texttt{a = Bool}) and instead create a
Boolean signal that is originally false and use the identity function to
not alter the hoverable information. With a simple change, we can create
a function that acts on \texttt{Signal Element} instead of
\texttt{Element}:

\begin{verbatim}
hoverablesJoin : Signal Element -> (Signal Element, Signal Bool)
hoverablesJoin elem =
    let pool = hoverables False
    in (lift (pool.hoverable id) elem, pool.events)
\end{verbatim}

Notice that \texttt{pool.hoverable} is partially applied to \texttt{id}
purely, and then lifted on to the argument. This is possible,
ultimately, because \texttt{pool.hoverable} is pure.

Just how novel is this small, but hugely significant change? The
\href{http://docs.elm-lang.org/library/Graphics/Input.elm\#hoverables}{documentation}
for \texttt{hoverables} states that it allows users to ``create and
destroy elements dynamically and still detect hover information,'' but
gives no further indicators on how to do so. Though the Elm website is
full of examples, there are none for either \texttt{hoverable} or
\texttt{hoverables}. Moreover, in the
\href{https://groups.google.com/d/msg/elm-discuss/QgowLy5jdhA/CZQfjkbjMsEJ}{mailing
list post} that introduced these functions, Czaplicki said that
``\texttt{hoverables} is very low level, but the idea is that you can
build any kind of nicer abstraction on top of it.'' We have done just
that.

When a element changes, \texttt{hoverablesJoin} attaches the same
hovering signal to the new element. If we were to display both the old
and the new elements on the screen, hovering over either would trigger
the signal.

More dangerously, it becomes easy to create an infinite loop. Suppose an
Element shrinks on hover. Suppose the cursor hovers on the Element,
which is then replaced by a smaller Element, so the cursor is no longer
hovering on the Element. Then the original Element is put back, but is
now being hovered on! This condition manifests itself as flickering
between the two Elements. It is unavoidable in any language or system
that allows hover targets to change size in response to hover events.

\subsection{Generalized Joins on \texttt{Graphics.Input}}

It is not just hover detection that follows the paradigm of two primitives
with singular and plural names. Most of Elm's wrappers around HTML GUI
components do so as well, and have analogous types. They are therefore receptive
of the same join technique. For text-labeled buttons, press events are
represented by unit or a provided identifier:

\begin{verbatim}
 button : String -> (Element, Signal ())
 buttons : a -> { events : Signal a,
                  button : a -> String -> Element }
\end{verbatim}

If we wish to create a button whose string label varies, the types allow
the same exact same technique to be applied. (We have not tested whether
this approach works, but the types allow it.) The same goes for text
fields, which can be given text based on what else is known in the
program, and checkboxes. As currently implemented, checkboxes have
Boolean state so dynamic behavior has limited use. However, we can
imagine the ability to disable (``gray-out'') a checkbox in response to
other choices in a form.

Drop-down menus do not use the same paradigm as the rest; there is
\texttt{dropDown} but no \texttt{dropDowns}. This is unfortunate not
only because there seems to be no way to supply a
\texttt{Signal {[}String{]}}, but because drop-downs are commonly used
in groups. Multiple menus reporting on one signal would be very
convenient. If the menus were dynamic, it would be useful for the API to
indicate when an event occurred because the user makes a selection vs.~when the
menu updated and the previously selected string was no longer available.

\texttt{Graphics.Input} does not currently support radio buttons, but one could
imagine them being added in a way similar to the others. However, it is worth
observing that the pattern itself leaves something to be desired, namely, that
the functions are impure. One can apply \texttt{hoverable} to an Element twice,
place the resulting Elements in in different places on the screen (or place one
not at all), and the signals of Booleans will be different. This is true,
analogously, of all the functions of similar type in the library. We know of no
alternative that has been proposed, and do not have one ourselves. However, we
point out this deficiency in order to illustrate that the library still has room
for incremental and perhaps even radical growth.

\section{Implementing Menus}\label{implementing-menus}

A menu can be thought of as an instance of a tree.

\texttt{data Tree a = Tree a [a]}

Using \texttt{hoverablesJoin}, we created a structure of type
\texttt{Tree (Signal Element, Signal Bool)} to represent our menus.
Using the combinator
\texttt{combine : {[}Signal a{]} -\textgreater{} Signal {[}a{]}}, we
were able to turn this structure into a
\texttt{Signal (Tree (Element, Bool))}. Once we had that structure, we
could render the menu using a pure function from
\texttt{Tree (Element, Bool) -\textgreater{} Element} and then lift it.
Similarly to limiting the use of the IO monad in Haskell code, it is
advantageous to limit the use of signals in Elm code.

Our first trick was to replace submenus not shown on the screen with
spacers. A spacer is simply a blank rectangle with a specified with and
height
(\texttt{spacer : Int -\textgreater{} Int -\textgreater{} Element}). For
every menu item, we create both the menu Element and a spacer. The
spacer has the same width as the Element, but a height of 1. We then
switch between menu and the spacer based on hover information. This
helped in two ways. First of all, it allowed us to choose between two
options without creating a signal of signals. Secondly, the spacer
helped to align the submenu with its parent element.

Our second trick fixed a race condition. A submenu was rendered to the
screen if one of three conditions was met: the mouse was hovering upon
its parent, the mouse was hovering upon it, or the mouse was hovering
upon any of its descendants. As the mouse moved from the parent to the
child, the child was often replaced with a spacer before it could detect
that the mouse was hovering over it. To fix this problem, we created the
following function:

\begin{verbatim}
extend : Signal Bool -> Signal Bool
extend b = lift2 (||) b (delay millisecond b)
\end{verbatim}

Using \texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
transition from True to False. By applying this function to every
hovering Boolean in the Menu structure, we gave the mouse time to move
from a menu Element to its submenu before that submenu disappeared.

Interestingly enough, we were able to implement menus without explicitly
storing the state of the menu. There are two ways to remember state in
Elm code: \texttt{foldp}, or Elm's implementation of Arrowized FRP known
as \texttt{Automaton}. However, in our implementation we use the current
browser DOM state to remember our state. Instead of storing the menu
currently being displayed on the screen, we can derive it based on which
Element (if any) the mouse is currently hovering over. The elimination
of state from our code allows it to be much cleaner.

Note that the input to our menu is a \texttt{Tree (Signal String)}, but
not a \texttt{Signal (Tree String)}. We are able to map from
\texttt{Signal Element -\textgreater{} (Signal Element, Signal Bool)},
but not from \texttt{Element -\textgreater{} (Element, Bool)}.
Therefore, if we tried to turn a \texttt{Signal (Tree String)} into a
\texttt{Signal (Tree (Element, Bool))}, we would need to lift an
impossible hoverable function of type
\texttt{Element -\textgreater{} (Element, Bool)} on each element.
Therefore, each individual menu element can contain dynamic information,
but the menu structure must be static. We can, in practice, get around
this restriction by creating a tree that is larger than necessary and
filling the unused nodes with empty string, which our implementation
handles appropriately.

A possible solution for that issue would be to write a function of type
\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}. Although the
\texttt{combine} combinator can turn a list of signals into a signal of
lists, there isn't an inverse to turn a signal of lists into a list of
signals. It is impossible to write the function
\texttt{split : Signal {[}a{]} -\textgreater{} {[}Signal a{]}} because
the size of the list in the input is dynamic, whereas the size of the
output list is static. Since the \texttt{Tree} data structure contains
lists, it is therefore impossible to write a function of type
\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}.

\section{Related Work: TodoFRP}\label{related-work-todofrp}

We examine \href{https://github.com/evancz/TodoFRP}{TodoFRP}, a simple
Elm web app created by Czapliki, to both pinpoint the problem and show
its universality. TodoFRP is the current state-of-the-art in highly
reactive Elm GUIs. It provides examples of different levels of
reactivity, a familiar context for Elm veterans, and a few dirty tricks
of its own.

TodoFRP presents the user with a text field asking, ``what needs to be
done?''. Entered TODO entries become DOM elements, which can be deleted
with a ``x'' button, also a DOM element. The button is implemented using
the \texttt{Graphics.Input} function

\begin{verbatim}
customButtons : a -> { events : Signal a,
                       customButton : a -> Element -> Element -> Element -> Element }
\end{verbatim}

Notice the similarity with \texttt{hoverables}. Each call of
\texttt{customButton} provides the identifier event when the button is
clicked, and three (pure) Elements to display: one normally, one on
hover, and one on click. The result is an \texttt{Element}, not a
\texttt{Signal Element}, that nevertheless changes among those three in
response to the mouse. This is possible because the result Element's
dimensions are taken to be the maximum of the three inputs' dimensions.
Even if the Elements have different sizes, the resulting Element and
therefore the hover surface remains fixed in size. Although the same
join technique can be applied, we find it likely that it would not work
as intended.
%Note to Max: I think it does work...
%We should try it out then.

In the case of TodoFRP, these Elements are different colors of the ``x''
and the same for each TODO entry. The polymorphic \texttt{a}s are unique
identifiers (ascending integers) for each entry.

The TODO label Elements are dynamic and do not detect hover information.
The button Elements that do detect hover information are known
statically. Without \texttt{hoverablesJoin}, it would be impossible to
dismiss a TODO by clicking on its text label.

\section{Conclusion: To the Elm
Community}\label{conclusion}

It's true that we've used the \texttt{hoverables} function in a way that
it was (probably) never meant to be used, and there are some caveats
involved in doing so. Many small GUIs do not require dynamic,
hover-detecting elements. However, most large mouse-based GUIs do, and
creating them in Elm will necessarily encounter the obstacles we have
described.

We have implemented all of this without language modifications. However,
it is hoped that as the community becomes more familiar with functional
GUIs, new libraries are added to incorporate some of our tricks, or even
make them unnecessary. Elm's upcoming third-party library sharing system
looks to be an excellent opportunity to refine abstractions and idioms
for GUIs.

We present these techniques and analysis in the hopes they are of
service to the Elm and FRP communities. We hope this paper becomes
useful in Elm's goal of making GUIs simpler to implement and more robust
to use.

\section*{Acknowledgments}
We would like to thank Evan Czaplicki and Stephen Chong for creating
Elm, and the Elm community for growing it. We thank Norman Ramsey for
his guidance through functional programming, and our paper reviewers,
\ldots{} .

%Make sure to put a link to our GitHub
\bibliographystyle{alpha}
\bibliography{refs.bib}
\end{document}